---
title: "20200202_bikeshareDemand"
author: Mike Gaunt
date: February 02, 2020
output:
  github_document:
    fig_width: 5
    fig_height: 5
    dev: jpeg

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## GitHub Documents

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

## Analysis Set-up

### Library Installation

```{r cars}
library(tidyverse)
library(tidymodels)
library(magrittr)
library(lubridate)
library(skimr)

library(ggplot2)
library(ggcorrplot)
library(janitor)
```

### Raw Data Inport
```{r}
bikeshare = readr::read_csv("./data/austin_bikeshare_trips.csv", 
                          col_names = TRUE) 

weather = read.csv("./data/austin_weather.csv", 
                        header = TRUE, stringsAsFactors = FALSE) %>%  
  janitor::clean_names() %>%  
  mutate(date = date %>%  
           as_date())
```

```{r}
bkshr = bikeshare %>% 
  mutate(date = start_time %>%
           as_date()) %>%
  group_by(date) %>% 
  summarise(trips = n()) %>%  
  merge(weather, by = "date", 
        all.x = T) %>%  
  mutate(across(!c(events, date), as.numeric), 
         events = case_when(events == " "~"no_event",
                            T~events) %>% 
           as.factor()) 
```

```{r}
bkshr %>%  
  skim()
```

Taking a look at PCA.
```{r}
bkshr_pca = bkshr %>%  
  select(date, where(is.numeric), -trips) %>%
  na.omit() 

bkshr_pca_rec = recipe(~., bkshr_pca) %>%
  step_normalize(all_numeric()) %>% 
  step_medianimpute(all_numeric()) %>%
  step_pca(all_numeric())

bkshr_pca_rec %>%  
  prep() %>%  
  juice()
  tidy(3)
tidy()

bkshr_pca_prep = bkshr_pca_rec %>%  
  prep() 

sdev = bkshr_pca_prep$steps[[1]]$sds
sdev = bkshr_pca_prep$steps[[2]]$res$sdev
percent_variation <- sdev^2 / sum(sdev^2)
var_df <- data.frame(PC=paste0("PC",1:length(sdev)),
                     var_explained=percent_variation,
                     stringsAsFactors = FALSE)

var_df %>%
  mutate(PC = fct_inorder(PC)) %>%
  ggplot(aes(x=PC,y=var_explained))+geom_col()

install.packages("zoo")
```

```{r}
names_glue_lag = "{.col}_lag_{.fn}" 
fns_lag = purrr::map(c(1,7, 14), ~purrr::partial(dplyr::lag, n = .x))
names_glue_rll = "{.col}_rll_{.fn}." 
fns_rll = purrr::map(c(3,5,7, 14), ~purrr::partial(zoo::rollmean, k = .x, align = "right", fill = T))

model_data = bkshr %>%  
  dplyr::mutate(dplyr::across(trips, fns_lag, .names = names_glue_lag), 
                dplyr::across(trips_lag_1 , fns_rll, .names = names_glue_rll)) %>%  
  select(date, starts_with("trips"), everything()) %>%  
  na.omit()
```



## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
yolo = model_data %>%  
  pivot_longer(cols = contains("trips")) %>% 
  ggplot() + 
  geom_line(aes(date, value, color = name))


```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.



## Create Dat Model 
```{r}
bkshr_split = initial_time_split(model_data, prop = .8)
bkshr_train = training(bkshr_split)
bkshr_test = testing(bkshr_split)
```

```{r}
folds = rsample::rolling_origin(bkshr_train, initial = 100)

folds$splits[[1]] %>%  
  analysis() 
  

folds$splits[[2]] %>%  
  analysis()
```

```{r}
model_rec = recipe(trips~., data = model_data) %>% 
  step_date(date, features = c("dow", "month", "year")) %>%  
  step_holiday(date) %>% 
  step_normalize(all_numeric(), -all_outcomes(), -contains("date")) %>% 
  step_corr(all_numeric(), -all_outcomes()) %>% 
  step_dummy(all_nominal()) %>%  
  step_nzv(all_predictors())
```


```{r}

mdl_mars = mars(
  num_terms = tune(),
  prod_degree = tune(),
  prune_method = "backward"
) %>% 
  set_engine("earth") %>% 
  set_mode("regression") %>% 
  translate()

mars_wrkfl = workflow() %>%  
  add_recipe(model_rec) %>%
  add_model(mdl_mars)

mars_wrkfl %>%  
  parameters() %>%  
  pull_dials_object("num_terms")
  prod_degree()


```




```{r}
grid = crossing(
  num_terms = c(5, 10, 15, 20),
  prod_degree = c(1, 2)
)

doParallel::registerDoParallel()



tictoc::tic()
# mars_train = mars_wrkfl %>%
#   fit_resamples(
#     resamples = folds,
#     control = control_resamples(save_pred = TRUE, extract = get_model)
#   )

mars_tune = mars_wrkfl %>%
  tune_grid(
    resamples = folds,
    grid = grid,
    control = control_resamples(save_pred = TRUE, extract = get_model)
  )
tictoc::toc()
tictoc::toc()

```


```{r}
mars_train %>%  collect_metrics()

  yolo = mars_train %>%  
  collect_predictions() %>%  
  pivot_longer(cols = c(.pred, trips)) %>% 
  ggplot(aes(.row, value, color = name)) + 
  geom_point() + 
  geom_line()
  
  plotly::ggplotly(yolo)
  
  mars_train %>%  
    show_best(metric = "rmse")
  
  
mars_train$.extracts[[550]][".extracts"] %>%  
  unnest()
mars_train[796,]

mars_tune %>% 
  saveRDS("./rds/20200202_bikeshareDemand_mars_trained")
  getwd()

  mars_tune %>%  
  collect_metrics() %>%  
  filter(.metric == "rmse") %>%  
  ggplot() + 
  geom_line(aes(num_terms, mean, color = as.factor(prod_degree))) + 
  ylim(0, NA)

mars_final = mars_tune %>%
  select_best() 

mars_wrkfl_final = mars_wrkfl %>%
  finalize_workflow(mars_final) 



mars_fit %>%  
  collect_metrics()

mars_fit %>%  
  pull(.predictions)

mars_fit %>%  
  fit(bkshr)

pull_workflow_fit(mars_fit)$fit
mars_wrkfl_final %>% 
fit(bkshr)

predict(final_model, new_data = new_woman)
```


```{r}
mars_fit = mars_wrkfl %>%  
  fit(bkshr_train)

mars_fit %>%  show_model_info()

yolo = bkshr_test %>%  
  bind_cols(mars_fit %>%  
  predict(bkshr_test)) %>%
  pivot_longer(cols = c(trips, .pred)) %>% 
  ggplot(aes(date, value, color = name)) + 
  geom_line() + 
  geom_point()

plotly::ggplotly(yolo)

bkshr_test %>%  
  bind_cols(mars_fit %>%  
  predict(bkshr_test)) %>%  
  mutate(error = trips-.pred) %>%
  ggplot() +
  geom_point(aes(date, error))
  rmse(truth = trips, estimate = .pred)
  
  bkshr_test %>%  
  bind_cols(mars_fit %>%  
  predict(bkshr_test)) %>%  
  mutate(error = trips-.pred) %>%
  ggplot() +
  geom_histogram(aes(error), bins = 60)


```






